---
title: "Pasta Pass Analysis"
output: 
  flexdashboard::flex_dashboard:
    logo: logo.png
    vertical_layout: scroll
    css: theme_bridge.css
---

```{r setup, include=FALSE}

#Load required packages
setwd("C:/Users/Bridget/Documents/PastaPass")
library(flexdashboard)
library(tidyverse)
library(plotly)
library(knitr)
library(kableExtra)

#Set custom ggplot theme
ogPalette <- c("#A7AE00","#9A3324","#552F1A","#E3E0DB","#EDECEA")

theme_bridge <- function (base_size = 12)
{
  theme_classic(base_size = base_size) +
    theme(axis.title.x = element_text(margin = margin(t=10)),
          axis.title.y = element_text(margin = margin(r=10)),
          axis.text.x=element_text(hjust=1,angle=45),
          axis.ticks.x=element_blank(),
          legend.text=element_text(size=rel(1.05)),
          legend.title=element_text(size=rel(1.1))
    )
}

#Custom functions to format data
as.money <- function(x) {
  paste0('$',formatC(as.numeric(x),format='f',digits=2,big.mark=','))
}

#Load and clean data
food <- read_csv("food.csv")

leftovers <- read_csv("leftovers.csv") %>% 
  mutate(Date_ID=Date_ID %>% as.POSIXct(format="%d-%b-%y",tz="EST"))

people <- read_csv("people.csv") %>% 
  mutate(Guest=if_else(Guest!="",TRUE,FALSE)) %>% 
  group_by(Visit_ID,Dom,Joe,Megan,Guest) %>% tally(Guest) %>% 
  mutate(Guest=if_else(is.na(Guest),FALSE,Guest),
         Num_Pals = Dom+Joe+Megan,
         Num_Guests = n,
         Party_Size = Num_Pals+Num_Guests+1) %>% 
  select(-n)

guests <- read_csv("people.csv") %>% 
  filter(!is.na(Guest)) %>% 
  mutate(Guest_Cost=Guest_Cost %>% as.character() %>% str_replace("[$]","") %>% as.double()) %>% 
  select(Visit_ID,Guest,Guest_Cost)

breadsticks <- read_csv("breadsticks.csv") %>% 
  mutate(Date_ID=Date_ID %>% as.POSIXct(format="%d-%b-%y",tz="EST"))

cost <- read_csv("cost.csv") %>%
  mutate(Cost=Cost %>% as.character() %>% str_replace("[$]","") %>% as.double(),
         Tip=Tip %>% as.character() %>% str_replace("[$]","") %>% as.double(),
         Extra=Extra %>% as.character() %>% str_replace("[$]","") %>% as.double())

dates <- read_csv("dates.csv") %>% 
  mutate(Date_ID=Date_ID %>% as.POSIXct(format="%d-%b-%y",tz="EST"),
         Arrival_Time=Arrival_Time %>% as.POSIXct(format="%r"),
         Food_Time=Food_Time %>% as.POSIXct(format="%r"),
         Leave_Time=Leave_Time %>% as.POSIXct(format="%r"),
         Wait_Time = (Food_Time - Arrival_Time) %>% as.integer(),
         Dine_Time = (Leave_Time - Arrival_Time) %>% as.integer())

expenses <- read_csv("expenses_clean.csv") %>%
  mutate(Date_ID = Date_ID %>% as.POSIXct(format="%m/%d/%Y"),
         krog_flag = if_else(cat=="GROCERY","KROGER",
                             if_else(desc=="OLIVE GARDEN","OLIVE GARDEN",
                                     if_else(cat=="FOOD","EATING OUT","OTHER (NON FOOD)"))))

```

Overview {data-orientation=rows}
===================================

Welcome {.sidebar data-width=300}
------------------------------

September 25, 2017 should have been a day like any other, but due to an impulsive decision made under high pressure, September 25 marked the start of a journey that would challenge my mental and physical limits and that would change my way of living for the following 8 weeks. That Monday was Day 1 of the Olive Garden Pasta Pass providing me, the lucky purchaser, with 56 days of unlimited never ending pasta bowls, soup or salad, and breadsticks.

I decided to make what seemed like a worthless purchase into a worthwhile endeavor, making notes and collecting many data points about each visit in order to share the experience with others and include some more productive activities into my 'extra-curricular' life. The first result of this resolve is a blog, 1 of 22,000: My Pasta Pass Journey (link disabled), which provided a creative outlet for my reflections on this unusual period in my life. The second result is this dashboard, which provided an opportunity to practice using ggplot and flexdashboard.

Please enjoy exploring the tabs and learning more about my Pasta Pass journey!

Row 1 {}
--------------------------------

### The Basics

I made a total of `r max(food$Visit_ID)` visits to Olive Garden ordering `r max(food$Bowl_ID)` bowls of pasta.
Taking leftovers into account, I ate Olive Garden `r max(food$Visit_ID) + leftovers %>% filter(Meal_Type != 'Freezer') %>% nrow()` times over 8 weeks. Excluding breakfast, that means pasta bowls made up `r round((max(food$Visit_ID) + leftovers %>% filter(Meal_Type != 'Freezer') %>% nrow())/(8*7*2)*100)`% of my meals.

Joe, Megan, and Dominick also had the pasta pass. While we dined together for many meals, I ventured on `r max(food$Visit_ID) - nrow(people)` solo dines. Additionally over the course of the 8 weeks, `r length(unique(guests$Guest))` guests joined in on the OG experience.

There were `r 7*6*5` possible pasta bowl combinations. While I tried each of the 7 pastas, 6 sauces, and 5 meat toppings, I only made it through `r food %>% select(Pasta,Sauce,Topping) %>% unique() %>% nrow()` unique combinations. All washed down with `r breadsticks %>% select(Bridget) %>% sum()` breadsticks. See the "Food Info" tab for more fun facts related to the meals eaten over the journey.

That many breadsticks really can't be good for a person in the long term can it? Was the 8 weeks of unlimited Olive Garden really worth it? The true cost of my meals at Olive Garden would have been `r sum(cost$Cost) %>% as.money()`. I spent `r '$100.00'`. That's `r (sum(cost$Cost) - 100) %>% as.money()` worth of free pasta! See the "Pass Economics" tab for how this stacks up to my usual food spend.

Finally, how much time did I actually spend at Olive Garden? For each visit, I recorded the time at which I entered the restaurant, recieved my pasta bowl, and left the restaurant. Over the 8 weeks, I spent `r sum(dates$Dine_Time,na.rm=T)` minutes in Olive Garden. (That's `r (sum(dates$Dine_Time,na.rm=T)/60) %>% round(1)` hours!) The average visit time was `r mean(dates$Dine_Time,na.rm=T) %>% round(0)` minutes and the average wait until we received our entrees was `r mean(dates$Wait_Time,na.rm=T) %>% round(0)` minutes. The plot below shows the length of each meal and wait time as well as the location and party size.


Row 2 {}
------------------------------

### Chronology of Pass

```{r chronology}
chron.plot <- left_join(dates,people) %>% ggplot(aes_string(x="Date_ID", y="Food_Time")) +
  geom_pointrange(aes(ymin = Arrival_Time, ymax = Leave_Time,
                      text = paste('Arrival: ',Arrival_Time %>% strftime(format="%l:%M %p",tz="UCT"),'\n',
                                   'Food: ',Food_Time %>% strftime(format="%l:%M %p",tz="UCT"),'\n',
                                   'Departure: ',Leave_Time %>% strftime(format="%l:%M %p",tz="UCT"),'\n',
                                   'Visit Length: ',Leave_Time - Arrival_Time,'minutes \n',
                                   'Location: ',Location,'\n',
                                   'Party Size: ',if_else(is.na(Party_Size),1,Party_Size))),
                  color="#a7ae00") +
  scale_y_datetime(name = "Time of Day", date_breaks = "1 hour", date_labels = "%l:%M %p") +
  scale_x_datetime(name = "Date", date_breaks = "1 day", date_labels = "%b %e") +
  theme_bridge()

ggplotly(chron.plot,tooltip='text') %>% 
  layout(margin=list(l=100,b=80))

```

Food Info {data-orientation=rows}
===================================

Row 1
------------------------------

### Breadsticks

```{r breadsticks}

breadsticks.long <- breadsticks %>% select(-Visit_ID) %>%
    gather(key="person",value="value",Bridget,Dom,Megan,Joe) %>% 
  group_by(Date_ID,person) %>% summarise(value=sum(value)) %>% 
  group_by(person) %>% mutate(cum_value = cumsum(value)) %>% 
  filter(value > 0) %>% 
  ungroup()

bs.plot <- breadsticks.long %>% ggplot(aes(x=Date_ID,y=cum_value,color=person,group=person,
                                          text=paste(person,'\n',
                                                    'Date: ',Date_ID %>% strftime(format="%b %e"),'\n',
                                                    'Daily Count: ',value,'\n',
                                                    'Cumulative Count: ',cum_value))) +
  geom_point() +
  geom_line() +
  scale_y_continuous(name = "Breadsticks", breaks = seq(0,100,5), 
                     sec.axis=sec_axis(~.*140,name="Calories", breaks = seq(0,14000,1000))) +
  scale_x_datetime(name = "Date", date_breaks = "1 day", date_labels = "%b %e") +
  theme_bridge() +
  theme(legend.title = element_blank(),
        legend.key.height=unit(5,units="pt"))

ggplotly(bs.plot,tooltip='text') %>% 
  layout(margin=list(l=100,b=80),
         legend=list(x=100,y=0.5))
```

Row 2 {data-height=400}
------------------------------

### Appetizers
```{r appetizers}

food %>% group_by(Appetizer) %>% summarise(Meals = n()) %>% arrange(Meals) %>% 
  kable("html") %>% 
  kable_styling(full_width = F,position="center")

```

### Pastas

```{r pasta}

pastapie <- ggplot(food) + 
  geom_bar(aes(x=factor(1), fill=Pasta),width=1) + 
  labs(x = NULL, y = NULL) +
  theme(axis.ticks=element_line(linetype=0),
        axis.text=element_blank(),
        panel.background=element_rect(fill="white"),
        legend.key.size=unit(20,units="pt"),
        legend.text=element_text(size=rel(1.05)),
        legend.title=element_text(size=rel(1.1))) +
  coord_polar(theta="y") +
  scale_fill_brewer(palette="BuPu")

pastapie
```

Row 3 {data-height=400}
------------------------------

### Toppings
```{r toppings}
topbar <- ggplot(food,aes(x=factor(1),fill=Topping)) +
  geom_bar() +
  scale_y_continuous(name="Bowls",limits=c(0,45),expand = c(0, 0)) +
  theme_bridge() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank()) +
  scale_fill_brewer(palette="BuPu")
           
topbar
```

### Sauces
```{r sauce,fig.width=8}
saucebar <- ggplot(food) +
  geom_bar(aes(x=Sauce)) +
  scale_y_continuous(name="Bowls",limits=c(0,20),expand = c(0, 0)) +
  theme_bridge() +
  theme(axis.title.x = element_blank())
           
saucebar
```

Pass Economics {data-orientation=rows}
===================================

Row 1
------------------------------

### Pre Pass Share of Wallet (SOW)

```{r sow_pre}
#Typical share of wallet going to each category before pasta pass
exppre <- expenses %>% 
  mutate(cat=if_else(cat=="FOOD","EATING OUT",cat)) %>% 
  filter(period=='8wks prior',cat!="PAY") %>% 
  group_by(cat) %>% 
  summarise(net = sum(amount)) %>% 
  mutate(prop = net/sum(net))

exppre.plot <- exppre %>% ggplot(aes(x=reorder(cat,net),y=prop,
                                     text=paste('SOW: ',round(prop*100,1),'%')))+
  geom_bar(stat="identity") +
  scale_x_discrete(name = "Category") +
  scale_y_continuous(name = "Share of Wallet",breaks=seq(0,.50,.1),expand=c(0,0)) +
  theme_bridge()

ggplotly(exppre.plot,tooltip='text') %>% 
  layout(margin=list(l=70,b=110))
```

### Impact of Pasta Pass on SOW

```{r sow_pass}
#How SOW changed pre and post, highlighting food and Kroger

sowdat <- expenses %>% 
  filter(cat!="PAY") %>% 
  group_by(period,krog_flag) %>% 
  summarise(net = sum(amount)) %>% 
  mutate(prop = net/sum(net))

sow.plot <- sowdat %>% ggplot(aes(x=period,y=prop,fill=krog_flag,
                                  text=paste(krog_flag,'\n','SOW: ',round(prop*100,1),'%'))) +
  geom_bar(stat="identity",position="fill") +
  labs(x="",y="Share of Wallet") +
  theme_bridge() +
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=0,hjust=0)) +
  scale_fill_manual(values=ogPalette) +
  scale_y_continuous(expand = c(0, 0))

ggplotly(sow.plot,tooltip='text') %>% 
  layout(margin=list(l=70),
         legend=list(x=100,y=0.5))
```

Row 2
------------------------------

### Pre Pass Share of Stomach (SOS)

```{r sos_pre}
#Typical share of stomach going to Kroger before pass

foodpre <- expenses %>% 
  filter(period=='8wks prior',cat=="FOOD"|cat=="GROCERY") %>% 
  group_by(desc) %>% 
  summarise(net = sum(amount)) %>% 
  mutate(prop = net/sum(net)) %>% 
  filter(prop > .02)

foodpre.plot <- foodpre %>% ggplot(aes(x=reorder(desc,net),y=prop,
                                       text=paste('SOS: ',round(prop*100,1),'%')))+
  geom_bar(stat="identity") +
  scale_x_discrete(name = "Retailer") +
  scale_y_continuous(name = "Share of Stomach",breaks=seq(0,.50,.1),expand=c(0,0)) +
  theme_bridge()

ggplotly(foodpre.plot,tooltip='text') %>% 
  layout(margin=list(l=70,b=110))
```

### Impact of Pasta Pass on SOS

```{r sos_pass}
#Looking at how SOS changed pre and post, highlighting food and Kroger

sosdat <- expenses %>% 
  filter(cat=="FOOD"|cat=="GROCERY") %>% 
  group_by(period,krog_flag) %>% 
  summarise(net = sum(amount)) %>% 
  mutate(prop = net/sum(net))

sosdat.plot <- sosdat %>% ggplot(aes(x=period,y=prop,fill=krog_flag,
                                     text=paste(krog_flag,'\n','SOS: ',round(prop*100,1),'%'))) +
  geom_bar(stat="identity",position="fill") +
  labs(x="",y="Share of Stomach") +
  theme_bridge() +
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=0,hjust=0)) +
    scale_fill_manual(values=ogPalette) +
  scale_y_continuous(expand = c(0, 0))

ggplotly(sosdat.plot,tooltip='text') %>% 
  layout(margin=list(l=70,b=110),
         legend=list(x=100,y=0.5))
```

Row 3
------------------------------

### The Pass Dramatically Cut My Food Spending

```{r food_change}

foodchange <- expenses %>% 
  filter(cat=="FOOD"|cat=="GROCERY") %>% 
  group_by(period) %>% 
  summarise(net = sum(amount)) %>% 
  mutate(prop = net/sum(net))

foodchange.plot <- foodchange %>% ggplot(aes(x=period,y=-net))+
  geom_bar(stat="identity",position="dodge") +
  geom_text(aes(label=paste('$',-net)),nudge_y=50) +
  labs(x="",y="Total Food Spend ($)") +
  theme_bridge() +
  theme(axis.text.x=element_text(angle=0,hjust=.5))

foodchange.plot

```

### Sorry Kroger

```{r krog_change}
krogchange <- expenses %>% 
  filter(desc=="KROGER"&cat=="GROCERY") %>% 
  group_by(period) %>% 
  summarise(net = sum(amount)) %>% 
  mutate(prop = net/sum(net))

krogchange.plot <- krogchange %>% ggplot(aes(x=period,y=-net))+
  geom_bar(stat="identity",position="dodge") +
  geom_text(aes(label=paste('$',-net)),nudge_y=40) +
  labs(x="",y="Total Kroger Spend ($)") +
  theme_bridge() +
  theme(axis.text.x=element_text(angle=0,hjust=.5))

krogchange.plot

```

### OG Might Just Be the Winner Here

```{r oginc}

#build table with data to plot
oginc <- tribble(
  ~period, ~type, ~net,
  "8wks prior", "my spend", 0,
  "8wks prior", "guest spend", 0,
  "pasta pass", "my spend", 100 + sum(cost$Extra),
  "pasta pass", "guest spend", sum(guests$Guest_Cost,na.rm=T)
  )

oginc.plot <- oginc %>% ggplot(aes(x=period,y=net,fill=type)) +
  geom_bar(stat="identity") +
  annotate("text", label = as.money(sum(guests$Guest_Cost,na.rm=T)+sum(cost$Extra)+100), x = 2, y = 380) +
  annotate("text",label = as.money(0), x = 1, y = 10) +
  labs(x="",y="Total OG Spend ($)",fill="") +
  theme_bridge() +
  theme(legend.title = element_blank(),
        axis.text.x=element_text(angle=0,hjust=.5)) +
  scale_fill_manual(values=c("dimgray","gray"))

oginc.plot

```